Staged Abstract Interpreters

### 2019-09-05 paper/10_intro.tex
space
intro

Abstract interpretation as a lattice-based approach to sound static analyses was
proposed by \citet{DBLP:conf/popl/CousotC77}. Based on the notion of Galois
connections \cite{CousotCousot79-1}, the analyzer can soundly approximate the
concrete program runtime behavior by computing fixed points on an abstract
domain. Despite the tremendous theoretical development of abstract
interpretation over the years, constructing artifacts and analyzers that perform
sound abstract interpretation for modern and expressive languages was considered
complicated for a long time.

@@

Based on the idea of abstracting abstract machines, more
pragmatically, several implementation strategies utilizing purely
functional programming techniques to build abstract interpreters have
emerged. Such techniques include monads, monad transformers
\cite{DBLP:journals/pacmpl/DaraisLNH17,
  Sergey:2013:MAI:2491956.2491979}, arrows
\cite{Keidel:2018:CSP:3243631.3236767}, extensible effects
\cite{Kiselyov:2015:FMM:2804302.2804319, Githubsemantic}, etc. The pure
approaches provide certain benefits, including that the abstract interpretation
artifacts can be built compositionally and modularly, e.g., by using monad
transformers. Therefore, the soundness of analysis can be proven with less
effort, whether by mechanized proofs \cite{Darais:2016:CGC:2951913.2951934} or
paper-based proofs \cite{Keidel:2018:CSP:3243631.3236767}.  Also, referential
transparency and equational reasoning allow programmers to confidently reason
about the correctness of their implementations.

>>>>

Abstract interpretation as a lattice-based approach to sound static analyses was
proposed by \citet{DBLP:conf/popl/CousotC77, CousotCousot79-1}. Based on the notion of Galois
connections, the analyzer can soundly approximate the
concrete program runtime behavior by computing fixed points on an abstract
domain. Despite the tremendous theoretical development of abstract
interpretation over the years, constructing artifacts and analyzers that perform
sound abstract interpretation for modern and expressive languages was considered
complicated for a long time.

@@

%Based on the idea of abstracting abstract machines, more
%pragmatically, several implementation strategies utilizing purely
%functional programming techniques to build abstract interpreters have
%emerged.

Based on the idea of AAM, more pragmatically, several implementation strategies
in high-level functional programming languages have emerged.  Such techniques
include monads and monad transformers \cite{DBLP:journals/pacmpl/DaraisLNH17,
Sergey:2013:MAI:2491956.2491979}, arrows
\cite{Keidel:2018:CSP:3243631.3236767}, extensible effects
\cite{Kiselyov:2015:FMM:2804302.2804319, Githubsemantic}, etc. The pure
approaches provide certain benefits, for examples, the abstract interpretation
artifacts can be built in a compositional and modular fashion, and referential
transparency allow programmers to confidently reason about the correctness of
their implementations.  The soundness of analysis can also be proven with less
effort, whether by mechanized proofs \cite{Darais:2016:CGC:2951913.2951934} or
paper-based proofs \cite{Keidel:2018:CSP:3243631.3236767}.
